---
title: "Predicting The Manner of Weight-Lifting Exercise Using Quantified Self Devices"
author: "Zecharias Chee"
date: "`14 MAy 2016"
output: html_document
---
  
#Executive Summary
In this report, a random forest model is built to predict the manner in which a group of 6 people perform their barbell-lifts based on the weight-lifting data from [Groupware](http://groupware.les.inf.puc-rio.br/har). The training data was partitioned into (60%) training set and (40%) cross-validation set. The training set was used to build the random forest model and the cross-validation set was used to evaluate the model performance and estimate the out-of-sample error. The model was then applied to the testing data once to get the final results.

#Preprocessing

The following code block will reading the data from local workspace (assuming that the files are downloaded and reside in the working directory).

```{r}
#Exploratory Analysis shows that some values are #DIV/0, this should be treated as missing values.
training <- read.csv("pml-training.csv", na.strings=c("NA", "#DIV/0!"))
testing <- read.csv("pml-testing.csv", na.strings=c("NA", "#DIV/0!"))

dim(training)

str(training)
```

Next, we partition the training data into training set (which is used to train the model) and CV set (which is used to estimate out-of-sample error).

```{r}
library(caret)

# Partition into training set 60% + CV set 40%
set.seed(12345)
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTrain <- training[inTrain,]
myCV <- training[-inTrain,]

dim(myTrain)
dim(myCV)
summary(myTrain)
```

Notice that there are a few variables with just one unique value, all NAs, or very imbalanced. These variables will have insignificant predictive power and we would remove such zero covariates by doing the following.

```{r}
nzv <- nearZeroVar(myTrain)
myTrain <- myTrain[, -nzv]
```

Furthermore, the first variable (id) should be omitted from the model because it doesn't mean anything.

```{r}
myTrain <- myTrain[, -1]
```

Some variables have high counts of NA's (`r 11529/11776`%) and therefore should be removed.

```{r}
#Get the variables without NAs
myTrain <- myTrain[,apply(myTrain, 2, function(x) !any(is.na(x)))]

dim(myTrain)
```

Now we are ready to train the random forest. This can be done simply in caret with the *train* function:
  
  ```{r, cache=TRUE}
set.seed(12345)
control <- trainControl(method="cv", number=5)
modFit <- train(classe ~ ., data=myTrain, method="rf", trControl=control)
pred <- predict(modFit, newdata=myCV)

#Confusion Matrix
confusionMatrix(pred, myCV$classe)
```

Thus, we estimated an out-of-sample error of

```{r}
confusionMatrix(pred, myCV$classe)$overall[1]
```

Finally, we predict on the testing data and write the answers to text files for submission.

```{r}
answers <- predict(modFit, testing)

answers

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)
```